An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-25 VLDB2026/main.tex
1

	\textit{\textbf{Build peak memory.}}
	\textcolor{orange}{As explained in the build peak memory analysis for attribute filtering, the build peak memory depends on the sum of the index size and the dataset size, as shown in Figure~\ref{fig:rangeFilter_build_memory_mb}. Range filtering also follows this pattern. The reason why PASE deviates from this trend has been discussed earlier, and is not repeated here to save space.}

@@

		\textit{\textbf{\textcolor{orange}{Search peak memory.}}}
		\textcolor{orange}{As for the search peak memory usage, we can find from the Figure ~\ref{fig:rangeFilter_search_memory_mb} that it is basically consistent with the build peak memory usage trend. As we analyzed in the section on attribute filtering.
		}

@@

	Due to space constraints, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. The experimental results show that across all datasets (YT-Audio-Real, WIT-Real, and Deep1M) and query ranges, WinFilter consistently delivers the best performance. This highlights its \textcolor{red}{robustness}, which performs ANN search over only relevant tree nodes and merges results efficiently, making it highly effective for static range query scenarios.

	\textcolor{red}{iRange and Faiss} offers a good balance between query efficiency and recall, demonstrating stable performance across both datasets and range sizes. In contrast, UNIFY experiences a notable drop in performance under narrow query ranges, suggesting that its HSIG structure—despite incorporating skip lists for pre-filtering—still has limitations in such scenarios. However, by leveraging HNSW and bitmap-based post-filtering, UNIFY performs competitively under broader range conditions, demonstrating strong scalability.

>>>>

	\textit{\textbf{Build peak memory.}}  
	\textcolor{orange}{As explained in the attribute filtering analysis, the peak memory usage during index construction generally depends on the combined size of the dataset and the index, as illustrated in Figure~\ref{fig:rangeFilter_build_memory_mb}. Range filtering adheres to this pattern, while PASE deviates due to reasons discussed earlier.}

	\textcolor{orange}{WinFilter also presents an exception:  its peak memory usage is also affected by the level of parallelism. Although the final index size remains relatively small, the peak memory usage increases significantly during both construction and search. During construction, this is caused by the parallel building of local indices for each tree node, which results in substantial temporary memory usage. Similarly, during search, WinFilter needs to load multiple local indices in parallel for different tree nodes, further contributing to high peak memory consumption.}

@@

		\textit{\textbf{\textcolor{orange}{Search peak memory.}}}  
		\textcolor{orange}{As shown in Figure~\ref{fig:rangeFilter_search_memory_mb}, the peak memory usage during search generally follows a similar trend to that observed during index construction. This observation is consistent with our earlier analysis in the attribute filtering section.}

@@

	Due to space constraints, Figure~\ref{fig:exp_8_2} reports results for only the largest ($2^{-2}$) and smallest ($2^{-8}$) range ratios. The experimental results show that across all datasets (YT-Audio-Real, WIT-Real, and Deep1M) and query ranges, WinFilter consistently delivers the best performance. This highlights its robustnes, which performs ANN search over only relevant tree nodes and merges results efficiently, making it highly effective for static range query scenarios.

	iRange offers a good balance between query efficiency and recall, demonstrating stable performance across both datasets and range sizes. In contrast, UNIFY experiences a notable drop in performance under narrow query ranges, suggesting that its HSIG structure—despite incorporating skip lists for pre-filtering—still has limitations in such scenarios. However, by leveraging HNSW and bitmap-based post-filtering, UNIFY performs competitively under broader range conditions, demonstrating strong scalability.
